---
title: "SpanishNR_Q2NegRaisingInferencing_analysis"
output: html_document
date: "2023-12-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library

```{r}
library(car)
library(lme4)
library(lmerTest)
library(MuMIn)  # for R2 goodness of fit 
library(emmeans)
library(ggeffects) # for post-hoc pairwise comparisons 
library(glmmTMB)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(ggplot2)
library(ggsignif)
library(ggh4x)  # for nested axis labeling
```


# Load in dfs

```{r}
## Load in Q2 results (no participants were excluded)
Q2NRInferencingResults_SpanishNR <- read.csv("results-Q2NR_Inferencing-SpanishNR-clean.csv", header=T)
```

# Investigate the data

## How many counts are there for each group? 
```{r}
## By group
Q2_obs_by_group <- Q2NRInferencingResults_SpanishNR %>%
  group_by(condition) %>%
  summarize(n = n())


## By participant by group
Q2_obs_by_participant_group <- Q2NRInferencingResults_SpanishNR %>%
  group_by(Participant_ID,condition) %>%
  summarize(n = n())
```


## Find NR reading means

```{r}
## Find mean NR reading score per condition 
tapply(as.numeric(Q2NRInferencingResults_SpanishNR$Q2_as_num), 
       list(Q2NRInferencingResults_SpanishNR$condition), 
       mean)

## Find mean NR reading score per condition per participant
tapply(as.numeric(Q2NRInferencingResults_SpanishNR$Q2_as_num),
       list(Q2NRInferencingResults_SpanishNR$Participant_ID,
            Q2NRInferencingResults_SpanishNR$condition), 
       mean)
```


**conditions**   **expected NR reading freq.**  **actual NR reading freq.**
NR+subj+NPI                  NR                             0.86
NR+subj                      NR                             0.90
NNR+subj                   Non-NR                           0.15
NR+ind+NPI                   NR?                            0.80
NR+ind                       NR                             0.92
NNR+ind                    Non-NR                           0.12


## Checking that the items are crossed
```{r}
table(Q2NRInferencingResults_SpanishNR$item_number,
      Q2NRInferencingResults_SpanishNR$sentence_type,
      Q2NRInferencingResults_SpanishNR$mood
      )
```


# Stats model

## Re-leveling the sentence type factor ordering for the stats base comparisons
```{r}
## New base comparison level: NR
Q2NRInferencingResults_SpanishNR$sentence_type <- relevel(factor(Q2NRInferencingResults_SpanishNR$sentence_type), ref = "NR")
```


## glmer 
```{r}
## df: Q2NRInferencingResults_SpanishNR
## DV: as.factor(Q2_response)
## IVs: mood:sentence_type (2x3)
## random effects: Participant_ID, item_number

Q2.glmer=glmer(as.factor(Q2_response)~mood*sentence_type
               +(1|Participant_ID),    # fails to converge with item as RE
               data=Q2NRInferencingResults_SpanishNR, family="binomial")


## Check whether main effects or interaction is significant
Anova(Q2.glmer, type = "III")  # There is only a main effect of sentence_type, no interaction, no effect of mood
```

**Results**: 
- mood                 0.4114  1     0.5213 NS 
- sentence_type      249.5049  2     <2e-16 ***
- mood:sentence_type   2.4624  2     0.2919 NS 

### Removing mood as a non-significant factor 
```{r}
## Remove non-significant factors: 
Q2_sentencetype.glmer=glmer(as.factor(Q2_response)~sentence_type
                        +(1|Participant_ID),      # still fails to converge with item as RE
                        data=Q2NRInferencingResults_SpanishNR, family="binomial")

## compare models
anova(Q2.glmer,Q2_sentencetype.glmer) # not significantly different, so we use the simpler model

## added random slopes
Q2_sentencetype_RSs.glmer=glmer(as.factor(Q2_response)~sentence_type
                        +(1+sentence_type|Participant_ID),
                        data=Q2NRInferencingResults_SpanishNR, family="binomial")  # fails to converge
```

### results of modeling
```{r}
### summarize effect of IV, no direction of the effect
summary(Q2_sentencetype.glmer)

### estimates for the intercept and slopes of IV
Anova(Q2_sentencetype.glmer, type = "III")

## R2 - goodness of fit
r.squaredGLMM(Q2_sentencetype.glmer)
AIC(Q2.glmer)
```

**Results**:
NNR <<<(p<0.001) NR+NPI <(p<0.001) NR

Since there's no interaction, then no subsetting or anything for further inspections 


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
