---
title: "SpanishNR_Q1Acceptability_analysis"
output: html_document
date: "2023-12-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library

```{r}
library(car)
library(lme4)
library(lmerTest)
library(MuMIn)  # for R2 goodness of fit 
library(emmeans)
library(ggeffects) # for post-hoc pairwise comparisons 
library(glmmTMB)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(ggplot2)
library(ggsignif)
library(ggh4x)
```

# Load in dfs

```{r}
## Load in Q1 results (no participants were excluded)
Q1AcceptabilityResults_SpanishNR <- read.csv("results-Q1Acceptability-SpanishNR-clean.csv", header=T)
```


# Investigate the data

## Check that the groups and items are working
```{r}
## Check that each participant sees each verb in each condition once 
table(Q1AcceptabilityResults_SpanishNR$verb, 
      Q1AcceptabilityResults_SpanishNR$condition, 
      Q1AcceptabilityResults_SpanishNR$Participant_ID
      )


## Check that each participant sees each NPI type with each verb once 
table(Q1AcceptabilityResults_SpanishNR$verb, 
      Q1AcceptabilityResults_SpanishNR$NPI_type, 
      Q1AcceptabilityResults_SpanishNR$Participant_ID
      )


## Check that each participant sees each item once 
table(Q1AcceptabilityResults_SpanishNR$item_number, 
      Q1AcceptabilityResults_SpanishNR$Participant_ID
      )
```

Groups and Items are working as expected


## Find overall acceptability means 


```{r}
## Find mean acceptability score per condition 
tapply(as.numeric(Q1AcceptabilityResults_SpanishNR$Value), 
       list(Q1AcceptabilityResults_SpanishNR$condition), 
       mean
       )
```

*conditions*  *expected grammaticality*  *actual mean grammaticality*
NR+subj+NPI     grammatical   (6?)                 5.5
NR+subj         grammatical   (6?)                 6.2
NNR+subj        grammatical   (6?)                 6.1
NR+ind+NPI      ungrammatical (2?)                 4.0
NR+ind          grammatical   (5?)                 5.1
NNR+ind         grammatical   (6?)                 5.6


```{r}
## Find mean acceptability score per participant per condition 
tapply(as.numeric(Q1AcceptabilityResults_SpanishNR$Value),
       list(Q1AcceptabilityResults_SpanishNR$Participant_ID,
            Q1AcceptabilityResults_SpanishNR$condition
            ), 
       mean
       )


## Find mean acceptability score per verb per condition 
tapply(as.numeric(Q1AcceptabilityResults_SpanishNR$Value),
       list(Q1AcceptabilityResults_SpanishNR$verb,
            Q1AcceptabilityResults_SpanishNR$condition
            ), 
       mean
       )


## Find mean acceptability score per verb per item 
tapply(as.numeric(Q1AcceptabilityResults_SpanishNR$Value),
       list(Q1AcceptabilityResults_SpanishNR$item_number,
            Q1AcceptabilityResults_SpanishNR$condition
            ), 
       mean
       )
```



# Stats modeling

## z-score the 1-7 likert scale responses 
```{r}
## z-score the Likert scale values in the df to be modeled 
Q1_z <- Q1AcceptabilityResults_SpanishNR
Q1_z$Value <- scale(as.numeric(Q1_z$Value))
```


## lmer 
```{r}
## df: Q1_z
## DV: Value
## IVs: mood:sentence_type 
## random effects: Participant_ID, item_number

Q1.lmer=lmer(as.numeric(Value)~mood*sentence_type
             +(1|Participant_ID)+(1|item_number), 
                     data=Q1_z)  # converges

Anova(Q1.lmer, type="III")  # everything is significant, nothing removed from model 
```

### results of modeling 
```{r}
### summarize effect of IV, no direction of the effect
Anova(Q1.lmer, type="III")

### estimates for the intercept and slopes of IV
summary(Q1.lmer)

## R2 - goodness of fit 
r.squaredGLMM(Q1.lmer)  # for R2c - percentage of variance of DV explained by model 
```

**Results** using lme4 version 1.1-26 & lmerTest version 3.1-3: 
ME: mood                19.0853  1  1.250e-05 ***
ME: sentence_type      204.7555  2  < 2.2e-16 ***
I:  mood:sentence_type  40.2132  2  1.853e-09 ***

*Please note that these χ2 results slightly differ from those presented at ELM3*


## Additional investigation: see if base model is improved by adding random slopes
```{r}
### Added mood and sentence type random slopes
Q1.RSs.lmer=lmer(as.numeric(Value)~mood*sentence_type
                 +(1+mood+sentence_type|Participant_ID)
                 +(1+mood|item_number),
                       data=Q1_z)       # most complex model that converges 

### compare with anova()
anova(Q1.lmer,Q1.RSs.lmer)  # Addition of random effects significantly improves model 
```

### results of modeling with random slopes
```{r}
### summarize effect of IV, no direction of the effect
Anova(Q1.RSs.lmer, type="III")

### estimates for the intercept and slopes of IV
summary(Q1.RSs.lmer)

## R2 - goodness of fit 
r.squaredGLMM(Q1.RSs.lmer)  # for R2c - percentage of variance of DV explained by model 
```

**Results**: Adding random slopes slightly changed results, in that the main effect of mood went from *** to **.
However, this doesn't change our analysis. 



## Pairwise comparisons using emmeans() 
```{r}
emmeans(Q1.lmer, pairwise ~ mood * sentence_type)
```

*Comparisons that we care about*:
- NR+ind vs NR+ind+NPI:        0.6452 0.0661 1632   9.755  <.0001  ***
- NR+subj vs NR+subj+NPI:      0.4195 0.0667 1635   6.292  <.0001  ***
- NR+ind+NPI vs NR+subj+NPI:  -0.8787 0.0667 1635 -13.178  <.0001  ***
- NNR+subj vs NR+subj:        -0.0871 0.0661 1632  -1.317  0.7759  NS


## Additional confirmation: Investigating interaction by subsetting

### First, subset by mood
```{r}
# First subset by mood
Q1ind_z <- subset(Q1_z, Q1_z$mood == "ind")
Q1subj_z <- subset(Q1_z, Q1_z$mood == "subj")
```

#### Modeling the indicatives
```{r}
# Model the IND subset 

## df: Q1ind_z
## DV: Value
## IVs: mood:sentence_type 
## random effects: Participant_ID, item_number

Q1ind.lmer=lmer(as.numeric(Value)~sentence_type
                +(1|Participant_ID)+(1|item_number), 
                     data=Q1ind_z)  # converges

Anova(Q1ind.lmer, type="III")  # sentence type is significant 


### added random slope
Q1ind.RSs.lmer=lmer(as.numeric(Value)~sentence_type
                    +(1+sentence_type|Participant_ID)
                    +(1+sentence_type|item_number),
                       data=Q1ind_z)  # most complex model that converges 
### compared models with anova()
anova(Q1ind.lmer,Q1ind.RSs.lmer)  # Adding sentence type as random slopes improves model
```

##### results of modeling 
```{r}
### summarize effect of IV, no direction of the effect
Anova(Q1ind.RSs.lmer, type="III")

### estimates for the intercept and slopes of IV
summary(Q1ind.RSs.lmer)

## R2 - goodness of fit 
r.squaredGLMM(Q1ind.RSs.lmer)  # for R2c - percentage of variance of DV explained by model 
```

**Results**: 
- sentence_type: 50.7934, p=9.34e-12 (***) 
NR+NPI << (p<.001) NR < (p=0.0129) NNR

#### Modeling the subjunctives
```{r}
# Model the SUBJ subset 

## df: Q1subj_z
## DV: Value
## IVs: mood:sentence_type 
## random effects: Participant_ID, item_number

Q1subj.lmer=lmer(as.numeric(Value)~sentence_type
                 +(1|Participant_ID)+(1|item_number), 
                     data=Q1subj_z)  # converges

Anova(Q1subj.lmer, type="III")  # is significant 


### added random slopes
Q1subj.RSs.lmer=lmer(as.numeric(Value)~sentence_type
                     +(1+sentence_type|Participant_ID)
                     +(1|item_number),
                       data=Q1subj_z)  # most complex model that converges 

### compared models with anova()
anova(Q1subj.lmer,Q1subj.RSs.lmer)  # Adding Sentence type as random slope improves model
```

##### results of modeling 
```{r}
### summarize effect of IV, no direction of the effect
Anova(Q1subj.RSs.lmer, type="III")

### estimates for the intercept and slopes of IV
summary(Q1subj.RSs.lmer)

## R2 - goodness of fit 
r.squaredGLMM(Q1subj.RSs.lmer)  # for R2c - percentage of variance of DV explained by model 
```

**Results**: 
- sentence_type: 35.674, p=1.793e-08 (***) 
NR+NPI << (p<.001) NR = (p=0.139814) NNR


### Second, subset by sentence type
```{r}
Q1_NNR_z <- subset(Q1_z, Q1_z$sentence_type == "NNR")
Q1_NR_z <- subset(Q1_z, Q1_z$sentence_type == "NR")
Q1_NRNPI_z <- subset(Q1_z, Q1_z$sentence_type == "NR+NPI")

```

#### Modeling the non-Neg-raising constructions
```{r}
# Model the NNR subset 

## df: Q1_NNR_z
## DV: Value
## IVs: mood
## random effects: Participant_ID, item_number

Q1_NNR.lmer=lmer(as.numeric(Value)~mood
                 +(1|Participant_ID)+(1|item_number), 
                     data=Q1_NNR_z)  # converges

Anova(Q1_NNR.lmer, type="III")  # mood is significant 


### added random slopes
Q1_NNR.RSs.lmer=lmer(as.numeric(Value)~mood
                     +(1+mood|Participant_ID)
                     +(1+mood|item_number),
                       data=Q1_NNR_z)  # most complex model that converges 

### compared models with anova()
anova(Q1_NNR.lmer,Q1_NNR.RSs.lmer)  # adding mood as random slopes improves model
```

##### results of modeling 
```{r}
### summarize effect of IV, no direction of the effect
Anova(Q1_NNR.RSs.lmer, type="III")

### estimates for the intercept and slopes of IV
summary(Q1_NNR.RSs.lmer)

## R2 - goodness of fit 
r.squaredGLMM(Q1_NNR.RSs.lmer)  # for R2c - percentage of variance of DV explained by model 
```

**Results**: 
- mood: 5.0116, p=0.02518 (*) 
IND < (p=0.0299) SUBJ


#### Modeling the plain Neg-raising constructions (without NPIs)
```{r}
# Model the NR subset 

## df: Q1_NR_z
## DV: Value
## IVs: mood
## random effects: Participant_ID, item_number

Q1_NR.lmer=lmer(as.numeric(Value)~mood
                +(1|Participant_ID)+(1|item_number), 
                     data=Q1_NR_z)  # converges

Anova(Q1_NR.lmer, type="III")  # is significant 


### added random slopes
Q1_NR.RSs.lmer=lmer(as.numeric(Value)~mood
                    +(1+mood|Participant_ID)
                    +(1|item_number),
                       data=Q1_NR_z)  # most complex model that converges 
### compared models with anova()
anova(Q1_NR.lmer,Q1_NR.RSs.lmer)  # adding mood as random slope improves model
```

##### results of modeling 
```{r}
### summarize effect of IV, no direction of the effect
Anova(Q1_NR.RSs.lmer, type="III")

### estimates for the intercept and slopes of IV
summary(Q1_NR.RSs.lmer)

## R2 - goodness of fit 
r.squaredGLMM(Q1_NR.RSs.lmer)  # for R2c - percentage of variance of DV explained by model 
```

**Results**: 
- mood: 42.0391, p=8.947e-11 (***) 
IND < (p<0.001) SUBJ


#### Modeling the Neg-raising constructions with NPIs
```{r}
# Model the NR+NPI subset 

## df: Q1_NRNPI_z
## DV: Value
## IVs: mood
## random effects: Participant_ID, item_number

Q1_NRNPI.lmer=lmer(as.numeric(Value)~mood
                   +(1|Participant_ID)+(1|item_number), 
                     data=Q1_NRNPI_z)  # converges

Anova(Q1_NRNPI.lmer, type="III")  # is significant 


### added random slopes to model
Q1_NRNPI.RSs.lmer=lmer(as.numeric(Value)~mood
                       +(1+mood|Participant_ID)+(1+mood|item_number),
                       data=Q1_NRNPI_z)  # most complex model that converges 

# compare both models with anova()
anova(Q1_NRNPI.lmer,Q1_NRNPI.RSs.lmer)  # adding mood as random slopes improves model
```

##### results of modeling 
```{r}
### summarize effect of IV, no direction of the effect
Anova(Q1_NRNPI.RSs.lmer, type="III")

### estimates for the intercept and slopes of IV
summary(Q1_NRNPI.RSs.lmer)

## R2 - goodness of fit 
r.squaredGLMM(Q1_NRNPI.RSs.lmer)  # for R2c - percentage of variance of DV explained by model 
```


**Results**: 
- mood: 49.435, p=2.051e-12 (***) 
IND < (p<0.001) SUBJ

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
